---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
# gender conscious routing

Routing along streets named after women rather than men.

<!-- badges: start -->
[![R build status](https://github.com/mpadge/gender-conscious-routing/workflows/R-CMD-check/badge.svg)](https://github.com/mpadge/gender-conscious-routing/actions?query=workflow%3AR-CMD-check)
[![codecov](https://codecov.io/gh/mpadge/gender-conscious-routing/branch/master/graph/badge.svg)](https://codecov.io/gh/mpadge/gender-conscious-routing)
[![Project Status: Concept](http://www.repostatus.org/badges/latest/concept.svg)](http://www.repostatus.org/#concept)
<!-- badges: end -->


## Gender catogorizer

This **R** package includes an internally bundled C library for categorising
the gender of first names, thanks to Michael Jörg (available
[here](https://www.heise.de/ct/ftp/07/17/182/)). The library is extremely fast
and flexible; covers all European languages and a host of others, and
categorises gender very accurately. Here's a test run on a very large data set
of names from the English-speaking world:

```{r load, echo = FALSE, message = FALSE}
devtools::load_all (".", export_all = FALSE)
```



```{r babynames}
u <- "https://github.com/hadley/data-baby-names/raw/master/baby-names.csv"
if (!file.exists ("baby-names.csv"))
    chk <- download.file (u, "baby-names.csv")
n <- read.csv ("baby-names.csv", stringsAsFactors = FALSE)
format (nrow (n), big.mark = ",")
st <- system.time (x <- get_gender (n$name))
st
knitr::kable (table (x$gender))
knitr::kable (table (n$sex))
```

Categorising `r format (nrow (n), big.mark = ",")` names took only `r st [3]`
seconds, or around 100,000 names per second. The following code compares the
accuracy, noting that many names are of course unisex, whereas the "baby-names"
data are direct records of individual names and sex. 

```{r babyname-output}
x$gender [x$gender == "IS_MALE"] <- "boy"
x$gender [x$gender == "IS_MOSTLY_MALE"] <- "boy"
x$gender [x$gender == "IS_FEMALE"] <- "girl"
x$gender [x$gender == "IS_MOSTLY_FEMALE"] <- "girl"

index_right <- which (x$gender == n$sex)
message (format (length (index_right), big.mark = ","), " / ",
         format (nrow (x), big.mark = ","),
         " of names correctly classified = ",
         formatC (100 * length (index_right) / nrow (x),
                  format = "f", digits = 1), "%")

```

Noting that the baby name records are structured over time, and include
many repeats of the same names, we can try to create "mostly girl/boy"
categories based on relative proportions.


```{r categorise-sex}
library (dplyr)

categorise_sex <-  function (sex, size) {
    # define relative proportions:
    # if > rel_props [2], then category is singular
    # else if > rel_props [1], then category is "mostly" singular,
    # else category is unisex
    rel_props <- c (4, 1000)

    if (length (size) == 1)
      return (sex)

    bi <- which (sex == "boy")
    gi <- which (sex == "girl")

    if (size [bi] > (size [gi] * rel_props [2]))
        return ("boy")
    else if (size [gi] > (size [bi] * rel_props [2]))
        return ("girl")
    else if (size [bi] > (size [gi] * rel_props [1]))
        return ("mostly boy")
    else if (size [gi] > (size [bi] * rel_props [1]))
        return ("mostly girl")
    else
        return ("unisex")
    }
n2 <- n %>%
    group_by (name, sex) %>%
    summarise (size = n ()) %>%
    group_by (name) %>%
    summarise (category = categorise_sex (sex, size))
knitr::kable (table (n2$category))
```

The above values for relative proportions were selected to give good agreement
with the observed overall distribution of categories as determined by the
internal library. These two more refined data sets can then be compared:

```{r}
n2$gender <- get_gender (n2$name)$gender
n2$gender [n2$gender == "IS_FEMALE"] <- "girl"
n2$gender [n2$gender == "IS_MALE"] <- "boy"
n2$gender [n2$gender == "IS_MOSTLY_FEMALE"] <- "mostly girl"
n2$gender [n2$gender == "IS_MOSTLY_MALE"] <- "mostly boy"
n2$gender [n2$gender == "IS_UNISEX_NAME"] <- "unisex"
```
Some names are simply not found, so we'll remove those from the comparison
before calculating final statistics.
```{r contingency}
n2 <- n2 [which (!n2$gender == "NAME_NOT_FOUND"), ]
knitr::kable (with (n2, table (category, gender)))
```
The accuracy in that case is
```{r contingency2}
ct <- with (n2, table (category, gender))
sum (diag (ct)) / sum (ct)
```
